---
title: "TA"
output: html_notebook
---



```{r}
# Library Load
library(tidyverse)
library(vtable)
library (lubridate)
library(scales)
library(patchwork)
```

```{r}

# loads csv into a base data frame
base_housing_data <- read_csv("fmhpi_master_file.csv")

# filters base data frame to build a data frame for only the Seattle-Tacoma-Bellevue WA area and between the years 2011 and 2019.  This "pre-covid" time frame starts when housing prices began to consistently rise 2011 in the Puget Sound area in 2011 and ends in Dec 2019. While the WHO did not declare COVID-19 a pandemic until March 2020, the effect of the disease on work, employment, etc. was already manifesting in February. 

base_puget_precovid_data <- filter(base_housing_data, GEO_Name == 
                            'Seattle-Tacoma-Bellevue WA') %>% 
    filter(Year >= 2011 & Year <= 2019)

# Creates a new YearMonth date variable as a date class (day is set to 01)
base_puget_precovid_data$YearMonth <- as.Date(with(base_puget_precovid_data,paste(Year,Month,1,sep="-")),"%Y-%m-%d") 
  
# Narrows the number of variables by removing the COde variables and then reorders the variables.
wip_puget_precovid_data <- base_puget_precovid_data %>% 
  select(YearMonth, Year, Month, Index_NSA, Index_SA)

# The following code plots the data to determine stationarity 

#Converts the data to 2 time series
index_nsa_ts <- ts(wip_puget_precovid_data$Index_NSA, frequency = 12, start = c(2011))
index_sa_ts <- ts(wip_puget_precovid_data$Index_SA, frequency = 12, start = c(2011))

# The following plots the not seasonally adjusted (nsa) index then assesses the acf and pacf.
plot(index_nsa_ts, main = 'Pre-Covid Freddie Mac Housing Price Index from 2011 - 2019 ', ylab = 'Housing Index', xlab = 'Year')

acf(index_nsa_ts)
pacf(index_nsa_ts)

# The nsa index for the pre-covid period is not stationary, the plot shows the the mean is time independent. The acf also show a declining relationship, another indicator of a non-stationary process. 

# # The following plots the seasonally adjusted (sa) index then assesses the acf and pacf.

plot(index_sa_ts, main = 'Pre-Covid Freddie Mac Seasonally 
     Adjusted Housing Price Index from 2011 - 2019 ', ylab = 'Housing Index', xlab = 'Year')

acf(index_sa_ts)
pacf(index_sa_ts)

# The sa index for the pre-covid period is not stationary, showing the same behavior as the nsa index.  Of note, the seasonality in the nsa plot is captured by a small repeating wave along the plot that is not seen in the sa plot.

```
The plots shows that the indexes are not stationary. The following section will make the plots covariance stationary.

```{r}

# The following code log diff's the sna na index to make them stationary.
log_diff_nsa_ts <- diff(log(index_nsa_ts))
log_diff_sa_ts <- diff(log(index_sa_ts))

# The following plots the sna na log diff time indexes.
mean(log_diff_nsa_ts)

ts.plot(log_diff_nsa_ts)
abline(h = '.0068')

mean(log_diff_sa_ts)

ts.plot(log_diff_sa_ts)
abline(h = '.0062')


```

## Observation: 
1. The not seasonally adjusted (nsa) index has a nice repeating cycle with a consistent reversion to the mean of .0068.  
2. The seasonally adjusted (sa) index has a high persistence above the mean (.0062) returning to the mean in 2018/2019.That corresponds with the greatest index decrease in the not seasonally adjusted index.  

```{r}
# The following builds the acf and pacf plots for the stationary time series plots.
acf(log_diff_nsa_ts)
pacf(log_diff_nsa_ts)

acf(log_diff_sa_ts)
pacf(log_diff_sa_ts)

```

## Observation:
1. The acf for both nsa, sa have a geometrically decaying collation indicative of a AR model. 
2. The nsa pacf also shows a geometrical relationship indicative of a MA model. 
3. The sa pacf does not have a uniform decay.  
4. Will use a ARMA model for the nsa.   

```{r}
```

